### 2023-05-22 appendix.tex
up

The well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section. Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward closed.
%		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.

% In the case \Safe and \Bad are both downward closed, they can both be finite.

@@

Since $\Bad \cap \Safe = \emptyset$, the subcase $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction since two upward closed sets have not an empty intersection. We then consider the case where $\Bad$ is as upward-closed
as it can be without intersecting with $\Safe$,
i.e. 
such that there exists an upward-closed set $U$
such that $\Bad = U \setminus \Safe$.

@@

\begin{proof}
Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$. Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm. 
Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$. % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
hence the resilience problem is decidable.
\end{proof}

@@

\begin{theorem}\label{k-up-up}
The % three %uniform 
$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
compatibility, $\Safe=\uparrow \Safe$
and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
%	and \Bad is upward closed or downward closed.
\end{theorem}

>>>>

The well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section. Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed.
%		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.

% In the case \Safe and \Bad are both downward-closed, they can both be finite.

@@

Since $\Bad \cap \Safe = \emptyset$, the subcase $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction since two upward-closed sets have not an empty intersection. We then consider the case where $\Bad$ is as upward-closed
as it can be without intersecting with $\Safe$,
i.e. 
such that there exists an upward-closed set $U$
such that $\Bad = U \setminus \Safe$.

@@

\begin{proof}
Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$. Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm. 
Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$. % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
hence the resilience problem is decidable.
\end{proof}

@@

\begin{theorem}\label{k-up-up}
The % three %uniform 
$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
compatibility, $\Safe=\uparrow \Safe$
and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
%	and \Bad is upward-closed or downward-closed.
\end{theorem}

### 2023-05-22 application.tex
up

We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient. The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
\mathieu{ne pas oublier de mettre une définition ici}
Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

@@

\mathieu{
VASS donc la base de $\uparrow post^*(s_0)$ est calculable. 
Intersection de $\Bad$ et $\uparrow \post^*(s_0)$ est upward closed donc ça se fait
					de décider si c'est  $\subseteq \pred^{\leq k}(\Safe)$
	(re: $k$-résilience décidable voir Prop. 34)
Problème: on pourrait avoir $\uparrow post^*(s_0) \neq  post^*(s_0)$.
Si $ \Bad \cap \uparrow \post^*(s_0) \subseteq  \pred^{\leq k}(\Safe)$ alors on a aussi
$ \Bad \cap  \post^*(s_0) \subseteq  \pred^{\leq k}(\Safe)$.
Si non, alors
notre procédure décisionnelle pour vérifier
ça
donne un témoin.
Le témoin, on peut vérifier si il est bien dans $\post^*(s_0)$
(c'est juste la question de est-ce que on peut l'atteindre depuis $s_0$
et l'accessibilité est décidable).
Si il est effectivement dans $\post^*(s_0)$ alors c'est bon on a aussi un contre exemple
qui prouve que
$ \Bad \cap  \post^*(s_0) \nsubseteq  \pred^{\leq k}(\Safe)$.
Si le témoin n'est pas dans $\post^*(s_0) \ldots$ alors
$\ldots$
j'imagine on peut reprendre la procédure décisionnelle sur 
$ \Bad \cap \uparrow \post^*(s_0)$
mais en excluant ce premier témoin,
et ou bien on a que c'est bon si on enlève ce témoins alors on a bien 
$\subseteq \pred^{\leq k}(\Safe)$
ou bien ... on obtiens un nouveau témoin ... pour lui aussi on peut
vérifier si il est effectivement dans $\post^*(s_0)$, et ainsi de suite.
Si jamais la $k$-résilience n'est pas vérifiée, on vas bien finir par
trouver un témoin que ça n'est pas vérifiée.
Si jamais la $k$-résilience est vérifiée par contre ...
$ \Bad \cap  \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
doit être fini, on a dit.
Est-ce qu'il y aurais un moyen de calculer ça plutôt ?
Par exemple,
on regarde
$ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$,
comme on a le $\uparrow$ 
devant $\uparrow \post^*(s_0)$
alors
deux choses se produisent,
d'abords les "trous" entre les éléments de $ \post^*(s_0)$ sont comblés.
Sauf que si un "trou" est un contre exemple et qu'on a qu'un nombre fini de
trou alors "on voit bien" que c'est un trou en testant que
ça appartiens ou non à $ \post^*(s_0)$ et "ça va" on peut s'en débarasser.
L'autre chose qui se produit, c'est que on a les droites qui vont vers plus l'infini...
Ça c'est plus difficile à gèrer parce que si
il y a qu'un nombre fini d'éléments dans $ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
et pas dans
$ \Bad \cap  \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
ça se fait,
mais si il y a un nombre infini d'éléments dans
$ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
potentiellement pas dans le plus petit,
alors c'est plus compliqué de faire notre procédure décisionnelle.
Il nous faudrait pouvoir décider si $\post^*(s_0)$ est fini ou non.
Décider si $\post^*(s_0)$ est borné ou non.
Est-ce que ça peut se réduire à un submarking problem ?
Il faudrait prendre le complémentaire peut-être ...}

>>>>

We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient. The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.

Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

@@

\mathieu{
VASS donc la base de $\uparrow post^*(s_0)$ est calculable. 
Intersection de $\Bad$ et $\uparrow \post^*(s_0)$ est upward-closed donc ça se fait
					de décider si c'est  $\subseteq \pred^{\leq k}(\Safe)$
	(re: $k$-résilience décidable voir Prop. 34)
Problème: on pourrait avoir $\uparrow post^*(s_0) \neq  post^*(s_0)$.
Si $ \Bad \cap \uparrow \post^*(s_0) \subseteq  \pred^{\leq k}(\Safe)$ alors on a aussi
$ \Bad \cap  \post^*(s_0) \subseteq  \pred^{\leq k}(\Safe)$.
Si non, alors
notre procédure décisionnelle pour vérifier
ça
donne un témoin.
Le témoin, on peut vérifier si il est bien dans $\post^*(s_0)$
(c'est juste la question de est-ce que on peut l'atteindre depuis $s_0$
et l'accessibilité est décidable).
Si il est effectivement dans $\post^*(s_0)$ alors c'est bon on a aussi un contre exemple
qui prouve que
$ \Bad \cap  \post^*(s_0) \nsubseteq  \pred^{\leq k}(\Safe)$.
Si le témoin n'est pas dans $\post^*(s_0) \ldots$ alors
$\ldots$
j'imagine on peut reprendre la procédure décisionnelle sur 
$ \Bad \cap \uparrow \post^*(s_0)$
mais en excluant ce premier témoin,
et ou bien on a que c'est bon si on enlève ce témoins alors on a bien 
$\subseteq \pred^{\leq k}(\Safe)$
ou bien ... on obtiens un nouveau témoin ... pour lui aussi on peut
vérifier si il est effectivement dans $\post^*(s_0)$, et ainsi de suite.
Si jamais la $k$-résilience n'est pas vérifiée, on vas bien finir par
trouver un témoin que ça n'est pas vérifiée.
Si jamais la $k$-résilience est vérifiée par contre ...
$ \Bad \cap  \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
doit être fini, on a dit.
Est-ce qu'il y aurais un moyen de calculer ça plutôt ?
Par exemple,
on regarde
$ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$,
comme on a le $\uparrow$ 
devant $\uparrow \post^*(s_0)$
alors
deux choses se produisent,
d'abords les "trous" entre les éléments de $ \post^*(s_0)$ sont comblés.
Sauf que si un "trou" est un contre exemple et qu'on a qu'un nombre fini de
trou alors "on voit bien" que c'est un trou en testant que
ça appartiens ou non à $ \post^*(s_0)$ et "ça va" on peut s'en débarasser.
L'autre chose qui se produit, c'est que on a les droites qui vont vers plus l'infini...
Ça c'est plus difficile à gèrer parce que si
il y a qu'un nombre fini d'éléments dans $ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
et pas dans
$ \Bad \cap  \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
ça se fait,
mais si il y a un nombre infini d'éléments dans
$ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
potentiellement pas dans le plus petit,
alors c'est plus compliqué de faire notre procédure décisionnelle.
Il nous faudrait pouvoir décider si $\post^*(s_0)$ est fini ou non.
Décider si $\post^*(s_0)$ est borné ou non.
Est-ce que ça peut se réduire à un submarking problem ?
Il faudrait prendre le complémentaire peut-être ...}

### 2023-05-22 definitions.tex
changed macro definnition position
up

\newcommand{\pred}{\textsf{pred}}
\newcommand{\post}{\textsf{post}}
% \renewcommand{\succ}{\textsf{Succ}}

\newcommand{\Bad}{\textsf{Bad}}
\newcommand{\Safe}{\textsf{Safe}}

@@

\noindent
Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo. In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.

In what follows, a downward closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement).
%		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
and an upward closed set $U$ is represented by its finite set of minimal elements. \\
% $M_U$ of its  satisfying $U=\uparrow M_U$.

>>>>

@@

\noindent
Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo. In such \emph{Finite Antichains} sets $(X,\leq)$, every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward-closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.

In what follows, a downward-closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement).
%		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
and an upward-closed set $U$ is represented by its finite set of minimal elements. \\
% $M_U$ of its  satisfying $U=\uparrow M_U$.

### 2023-05-22 resilience.tex
up

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.

@@

In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward closed and $\Safe$ is upward closed.
%		
We deduce 
% \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.}
% \mathieu{changer ça}
decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed as seen in Section~\ref{synthesis}.

@@

% \mathieu{En fait de plus en plus j'ai l'impression que on peut se passer des neutres: même dans une situation où on aurait des configurations neutre j'aurais tendance à dire qu'on peut se ramener à un cas safe complémentaire de bad en duplicants les neutres en deux parties, une avec "on a vu un very bad récement il faut aller vers safe" (et qui dans le nouveau modèle seraient donc bad) et une "on a pas vu de very bad tout vas bien" (et qui dans le nouveau modèle seraient donc safe)
% $=>$ Rédiger ça plus proprement}
While considering a sequence of configurations, every prefix is always 
either in the case that the latest visited $\Bad$ configuration has already been followed by a $\Safe$ configuration
or in the case that it still has to. 
Thus, in context, there are no true ``neutral'' configurations: either a neutral configuration is contextually ``$\Safe$'' because it follows a visit in a $\Safe$ configuration, either it is contextually ``$\Bad$'' because it follows a visit in a $\Bad$ configuration and has, accordingly, to reach a $\Safe$ one.
Hence we will consider not only $\Bad$ and $\Safe$ disjoint but
% w.l.o.g.
	also
$\Bad$ complement of $\Safe$.

>>>>

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$.

@@

In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
%		
We deduce 
% \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed sets.}
% \mathieu{changer ça}
decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.

@@

% \mathieu{En fait de plus en plus j'ai l'impression que on peut se passer des neutres: même dans une situation où on aurait des configurations neutre j'aurais tendance à dire qu'on peut se ramener à un cas safe complémentaire de bad en duplicants les neutres en deux parties, une avec "on a vu un very bad récement il faut aller vers safe" (et qui dans le nouveau modèle seraient donc bad) et une "on a pas vu de very bad tout vas bien" (et qui dans le nouveau modèle seraient donc safe)
% $=>$ Rédiger ça plus proprement}
While considering a sequence of configurations, every prefix is always 
either in the case that the latest visited $\Bad$ configuration has already been followed by a $\Safe$ configuration
or in the case that it still has to. 
Thus, in context, there are no true ``neutral'' configurations: either a neutral configuration is contextually ``$\Safe$'' because it follows a visit in a $\Safe$ configuration, either it is contextually ``$\Bad$'' because it follows a visit in a $\Bad$ configuration and has, accordingly, to reach a $\Safe$ one.
Hence we will consider not only $\Bad$ and $\Safe$ disjoint but
% w.l.o.g.
	also
$\Bad$ complement of $\Safe$,
in which case
if $\Bad$ is downward-closed then $\Safe$ is upward-closed and vice-versa.
Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. they are both downward
closed or both upward-closed,
have also been studied, 
see 
Appendix~\ref{case down down}
and
Appendix~\ref{case up up} for results.
